A quantitative framework for investigating the reliability of empirical network construction: supplemental information

### 2018-07-04 manuscript/CI_interactions.tex
final updates
shrank Figure 1 to compile on manuscript central. Added table of PhyloPic image attributions

    \item  Descriptions of ecological networks typically assume that the same interspecific interactions occur each time a community is observed. This contrasts with the known stochasticity of ecological communities: community composition, species abundances, and link structure all vary in space and time. Moreover, finite sampling generates variation in the set of interactions actually observed. For interactions that have not been observed, most data sets will not contain enough information for the ecologist to be confident that unobserved interactions truly did not occur.
    \item Here we develop the conceptual and analytical tools needed to capture uncertainty in the estimation of pairwise interactions. To define the problem, we identify the different contributions to the uncertainty of an interaction. We then outline a framework to quantify the uncertainty around each interaction by combining data on observed co-occurrences with prior knowledge. We illustrate this framework using perhaps the most extensively sampled network to date. 
    \item We found significant uncertainty in estimates for the probability of most pairwise interactions. This uncertainty can, however, be constrained with informative priors. This uncertainty scaled up to summary measures of network structure such as connectance or nestedness. By using a comprehensively sampled network as a test case, we find that even with informative priors we are likely to miss many interactions that may occur rarely or under different local conditions. 
    \item Overall, we demonstrate the utility of our approach, given the importance of acknowledging the uncertainty inherent in network studies, and the utility of treating interactions as probabilities in pinpointing areas where more study is needed. Most importantly, we stress that networks are best thought of as systems constructed from random variables, the stochastic nature of which must be acknowledged for an accurate representation. Doing so will fundamentally change networks analyses and yield greater realism.
  % Can be up to 350 words for MEE. This is 290 exactly.
\end{enumerate}

@@

  \begin{figure}[h!]
    \caption{Three nested levels of uncertainty mean that an observed interaction matrix is unlikely to capture all of the interactions that can really occur at a site. Interaction uncertainty means that some interactions that truly occur (black squares in True matrix) are less likely than others (grey squares; lighter grey is less likely) based on species traits. For example, lions are less likely to predate upon elephants than zebras. If our trait model is incomplete (e.g., if we neglect group hunting in lions), we might assume that elephants are too large to ever be prey for lions and assign this interaction a probability of 0.\\
    \indent Process uncertainty describes the fact that interactions occur with different probabilities in any given sample. We show six example networks representing interactions observed over two days at three sites. Interactions which occur during fewer sampling days are less likely to be observed.\\
    \indent Detection uncertainty reflects the fact that not all interactions are equally detectable. For example, interactions involving cryptic species are less likely to be seen.\\
    \indent After applying these three layers of uncertainty, some interactions which truly occur are very unlikely to be observed. The observed matrix (lower right) is a subsample of the true matrix, with the probability of observing each interaction depending on the interaction, process, and detection uncertainty associated with it. Note that some low-probability interactions (e.g., lion predation on zebra) are included while others (e.g., lion predation on elephants) are not. For an illustration of the way in which these levels of uncertainty combine in simulated data, and the effects of uncertainty on network structure, see \emph{Appendix S1}.}
    \label{conceptual_fig}
    \begin{center}
    % \includegraphics*[height=.5\textheight]{figures/conceptual_fig.eps}
    \end{center}
    \end{figure}

  \begin{figure}[h!]
      \caption{Despite the replication in our empirical dataset, most galler-natural enemy pairs were never observed co-occurring and those that did co-occur rarely interacted. \textbf{A)} Here we show a histogram of the number of pairs of species observed co-occurring at least once. 9794 galler-enemy pairs were never observed co-occurring (not shown).
      \textbf{B)} Here we show a histogram of the number of observed interactions within pairs of co-occurring species. Species which co-occurred but never interacted are included. 
      \textbf{C)} Here we show, for each species pair, the number of observed interactions plotted against the number of observed co-occurrences. The red, dashed line indicates a 1:1 relationship between interactions and co-occurrences.}
      \label{histograms}
      \begin{center}
      % \includegraphics*[height=.6\textheight]{figures/GP_histogram.eps}
      \end{center}
      \end{figure}

  \begin{figure}[h!]
    \caption{Imperfect detection of interactions increases the number of samples required to be confident that an interaction never occurs. Assume that two species cannot interact (i.e., $\lambda=0$) and that the number of observed interactions follows a binomial distribution. Here we show the upper bound (solid black line) of a 95\% Clopper-Pearson true credible interval for the interaction probability $\lambda$ for a pair of species that has been observed co-occurring $n$ times but never interacting ($k = 0$). The upper limit of the credible interval only reaches 0.1 (dashed, red line) with 35 observations. Thus, adding more observations is useful in controlling uncertainty, but the number of observations added must be very high. }
    \label{upper_limits}
    \begin{center}
    % \includegraphics*[width=.8\textwidth]{figures/upper_limit_DG.eps}
    \end{center}
    \end{figure}

  \begin{figure}[h!]
    \caption{Using prior distributions based on a single site (Zillis in Graub\"{u}nden, Switzerland) from~\citet{Kopelke2017}, we calculate posterior distributions for the probability of interaction ($\lambda$) between two species that have not yet been observed interacting ($k = 0$). 
    Posterior distributions (curves) and 95\% credible intervals (lines at top of panel) for $\lambda$ narrow and approach zero as the number of observed co-occurrences ($n$) increases. Diamonds indicate the maximum likelihood estimator for the mean probability of interaction.
    Dashed lines indicate threshold probabilities of 0.01, 0.05, and 0.1. The number of samples required to obtain a 95\% credible interval below each threshold increases rapidly. It takes just over 20 observed co-occurrences to be 95\% confident that $\lambda<0.10$, approximately 50 co-occurrences to be 95\% confident that $\lambda<0.05$, and over 100 co-occurrences to be 95\% confident that $\lambda<0.01$.}
    \label{Salix_pdfs_cdfs}
    \begin{center}
    % \includegraphics*[width=.8\textwidth]{figures/GP_pdfs_increasing_N_Zillis.eps}
    \end{center}
    \end{figure}

  \begin{figure}[h!]
    \caption{Mean connectance, links per galler, nestedness (NODF), and links per natural enemy for networks assembled using posterior distributions based on a single site (Zillis in Graub\"{u}nden, Switzerland) from~\citet{Kopelke2017} occupied a narrow range. To obtain distributions of network properties, we created 100 ``posterior-sampling" networks and then, for each of these, created 100 ``detection-filter" networks by randomly sampling 50\%-99\% of the interactions included in the posterior-sampling network. This simulates imperfect detection of interactions in the field. Each point represents the mean (+/- SD) network property (e.g., connectance) obtained from a set of 100 detection-filter networks. The detection-filter networks cover a broader range of network properties than the posterior-sampling networks but the value of each property decreases as the strength of the detection filter increases. Values in the original web are indicated by dashed lines.}
    \label{posterior_webs}    
    \begin{center}
    % \includegraphics[width=.8\textwidth]{figures/GP_posterior_properties_Zillis.eps}
    \end{center}
    \end{figure}

>>>>

    \item  Descriptions of ecological networks typically assume that the same interspecific interactions occur each time a community is observed. This contrasts with the known stochasticity of ecological communities: community composition, species abundances, and link structure all vary in space and time. Moreover, finite sampling generates variation in the set of interactions actually observed. For interactions that have not been observed, most data sets will not contain enough information for the ecologist to be confident that unobserved interactions truly did not occur.
    \item Here we develop the conceptual and analytical tools needed to capture uncertainty in the estimation of pairwise interactions. To define the problem, we identify the different contributions to the uncertainty of an interaction. We then outline a framework to quantify the uncertainty around each interaction by combining data on observed co-occurrences with prior knowledge. We illustrate this framework using perhaps the most extensively sampled network to date. 
    \item We found significant uncertainty in estimates for the probability of most pairwise interactions. This uncertainty can, however, be constrained with informative priors. This uncertainty scaled up to summary measures of network structure such as connectance or nestedness. By using a comprehensively sampled network as a test case, we find that even with informative priors we are likely to miss many interactions that may occur rarely or under different local conditions. 
    \item Overall, we demonstrate the utility of our approach, given the importance of acknowledging the uncertainty inherent in network studies, and the utility of treating interactions as probabilities in pinpointing areas where more study is needed. Most importantly, we stress that networks are best thought of as systems constructed from random variables, the stochastic nature of which must be acknowledged for an accurate representation. Doing so will fundamentally change networks analyses and yield greater realism.
  % Can be up to 350 words for MEE.
\end{enumerate}

@@

  \begin{figure}[h!]
    \caption{Three nested levels of uncertainty mean that an observed interaction matrix is unlikely to capture all of the interactions that can really occur at a site. Interaction uncertainty means that some interactions that truly occur (black squares in True matrix) are less likely than others (grey squares; lighter grey is less likely) based on species traits. For example, lions are less likely to predate upon elephants than zebras. If our trait model is incomplete (e.g., if we neglect group hunting in lions), we might assume that elephants are too large to ever be prey for lions and assign this interaction a probability of 0.\\
    \indent Process uncertainty describes the fact that interactions occur with different probabilities in any given sample. We show six example networks representing interactions observed over two days at three sites. Interactions which occur during fewer sampling days are less likely to be observed.\\
    \indent Detection uncertainty reflects the fact that not all interactions are equally detectable. For example, interactions involving cryptic species are less likely to be seen.\\
    \indent After applying these three layers of uncertainty, some interactions which truly occur are very unlikely to be observed. The observed matrix (lower right) is a subsample of the true matrix, with the probability of observing each interaction depending on the interaction, process, and detection uncertainty associated with it. Note that some low-probability interactions (e.g., lion predation on zebra) are included while others (e.g., lion predation on elephants) are not. For an illustration of the way in which these levels of uncertainty combine in simulated data, and the effects of uncertainty on network structure, see \emph{Appendix S1}. Attributions for PhyloPic silhouette images are given in \emph{Appendix S1-A}.}
    \label{conceptual_fig}
    \begin{center}
    % \includegraphics*[height=.5\textheight]{figures/conceptual_fig.eps}
    % \includegraphics*[height=.5\textheight]{Figure_1.eps}
    \end{center}
    \end{figure}

  \begin{figure}[h!]
      \caption{Despite the replication in our empirical dataset, most galler-natural enemy pairs were never observed co-occurring and those that did co-occur rarely interacted. \textbf{A)} Here we show a histogram of the number of pairs of species observed co-occurring at least once. 9794 galler-enemy pairs were never observed co-occurring (not shown).
      \textbf{B)} Here we show a histogram of the number of observed interactions within pairs of co-occurring species. Species which co-occurred but never interacted are included. 
      \textbf{C)} Here we show, for each species pair, the number of observed interactions plotted against the number of observed co-occurrences. The red, dashed line indicates a 1:1 relationship between interactions and co-occurrences.}
      \label{histograms}
      \begin{center}
      % \includegraphics*[height=.6\textheight]{figures/GP_histogram.eps}
      % \includegraphics*[height=.6\textheight]{Figure_2.eps}
      \end{center}
      \end{figure}

  \begin{figure}[h!]
    \caption{Imperfect detection of interactions increases the number of samples required to be confident that an interaction never occurs. Assume that two species cannot interact (i.e., $\lambda=0$) and that the number of observed interactions follows a binomial distribution. Here we show the upper bound (solid black line) of a 95\% Clopper-Pearson true credible interval for the interaction probability $\lambda$ for a pair of species that has been observed co-occurring $n$ times but never interacting ($k = 0$). The upper limit of the credible interval only reaches 0.1 (dashed, red line) with 35 observations. Thus, adding more observations is useful in controlling uncertainty, but the number of observations added must be very high. }
    \label{upper_limits}
    \begin{center}
    % \includegraphics*[width=.8\textwidth]{figures/upper_limit_DG.eps}
    % \includegraphics*[width=.8\textwidth]{Figure_3.eps}
    \end{center}
    \end{figure}

  \begin{figure}[h!]
    \caption{Using prior distributions based on a single site (Zillis in Graub\"{u}nden, Switzerland) from~\citet{Kopelke2017}, we calculate posterior distributions for the probability of interaction ($\lambda$) between two species that have not yet been observed interacting ($k = 0$). 
    Posterior distributions (curves) and 95\% credible intervals (lines at top of panel) for $\lambda$ narrow and approach zero as the number of observed co-occurrences ($n$) increases. Diamonds indicate the maximum likelihood estimator for the mean probability of interaction.
    Dashed lines indicate threshold probabilities of 0.01, 0.05, and 0.1. The number of samples required to obtain a 95\% credible interval below each threshold increases rapidly. It takes just over 20 observed co-occurrences to be 95\% confident that $\lambda<0.10$, approximately 50 co-occurrences to be 95\% confident that $\lambda<0.05$, and over 100 co-occurrences to be 95\% confident that $\lambda<0.01$.}
    \label{Salix_pdfs_cdfs}
    \begin{center}
    % \includegraphics*[width=.8\textwidth]{figures/GP_pdfs_increasing_N_Zillis.eps}
    % \includegraphics*[width=.8\textwidth]{Figure_4.eps}
    \end{center}
    \end{figure}

  \begin{figure}[h!]
    \caption{Mean connectance, links per galler, nestedness (NODF), and links per natural enemy for networks assembled using posterior distributions based on a single site (Zillis in Graub\"{u}nden, Switzerland) from~\citet{Kopelke2017} occupied a narrow range. To obtain distributions of network properties, we created 100 ``posterior-sampling" networks and then, for each of these, created 100 ``detection-filter" networks by randomly sampling 50\%-99\% of the interactions included in the posterior-sampling network. This simulates imperfect detection of interactions in the field. Each point represents the mean (+/- SD) network property (e.g., connectance) obtained from a set of 100 detection-filter networks. The detection-filter networks cover a broader range of network properties than the posterior-sampling networks but the value of each property decreases as the strength of the detection filter increases. Values in the original web are indicated by dashed lines.}
    \label{posterior_webs}    
    \begin{center}
    % \includegraphics[width=.8\textwidth]{figures/GP_posterior_properties_Zillis.eps}
    % \includegraphics[width=.8\textwidth]{Figure_5.eps}
    \end{center}
    \end{figure}

### 2018-07-04 manuscript/supplemental.tex
shrank Figure 1 to compile on manuscript central. Added table of PhyloPic image attributions

>>>>

\section*{Appendix S1-A: Attributions for PhyloPic images used in Figure 1 (main text)}

  \begin{table}
  \caption{Attributions for silhouette images taken from PhyloPic.org and used in Figure 1 (\emph{main text}). Terms for the CC BY-NC-SA 3.0 license can be found at \emph{https://creativecommons.org/licenses/by-nc-sa/3.0/}. Terms for the CC BY 3.0 license can be found at \emph{https://creativecommons.org/licenses/by/3.0/}.}
  \begin{tabular}{l|ll}
  Image & PhyloPic taxon & Attribution \\
  \hline
  Lion & Panthera leo & Public domain (uncredited) \\
  Leopard & Panthera pardus & Lukasiniho (CC BY-NC-SA 3.0; image not changed) \\
  Jackal & Canidae & Brian Gratwicke (photo) and T. Michael Keesey (vectorization). (CC BY 3.0; image not changed) \\
  Elephant & Loxodonta africana & Jan A. Venter, Herbert H. T. Prins, David A. Balfour & Rob Slotow (vectorized by T. Michael Keesey) (CC BY 3.0; image not changed) \\
  Zebra & Equus quagga & Cathy (CC BY-NC-SA 3.0; image not changed) \\
  Impala & Aepyceros melampus & Public domain (unknown author) \\
  Hare & Leporidae & Sarah Werning (CC A 3.0; image not changed) \\
  Acacia & Acacia tree & Mattia Menchetti (public domain) \\
  Grass & \multicolumn{2}{c}{Not sourced from PhyloPic. Public domain.} \\
  \hline
  \end{tabular}
  \end{table}
